---
title: "Coming Soon: Lecture 1"
description: |
  Getting started!
draft: false
author:
  - name: Jannik Buhr 
    url: https://jmbuhr.de
    orcid_id: 0000-0002-9350-2629
date: 11-02-2020
output:
  distill::distill_article:
    self_contained: false
categories:
  - lecture
bibliography: ["../../references.bib"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
)
```

> # Disclaimer:
>
> This is not the finished lecture, but I wanted to give you a sneak peak of how the website will look.
> I am very much developing this in the open.

# Video

Watch today's video here:

<iframe width="560" height="315" src="https://www.youtube.com/embed/AgpWX18dby4" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen>

</iframe>

# Slides

Hover over the slides and press **f** for full screen.
Press **?** for a list of keyboard shortcuts.

```{r slides, echo=FALSE}
knitr::include_url("example_slides.html")
```

# Script

## Roadmap

Day 1

- installing R and RStudio
- calculator
- datatypes
- variables
- functions
- vectors
- lists
- dataframes
- grammar of graphics
- ggplot (Palmer penguins)
- exercises
  - 

Day 2

- RStudio tour
  - settings
  - appearance
  - projects (and why we don't use `setwd`)
  - folder structure
- Reading data with `readr`
  - https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-10-29
  - from file
  - from url
  - download file with R
- Workflow
  - https://jmbuhr.de/books/dataIntro19Master/img/workflow.png
- various states of messiness in data
  - leading lines, comments,
  - other formats
    - csv2, excel, tsv
  - import text data with RStudio
    - always copy the code!
- `dplyr` verbs
  - select
  - filter
  - mutate
    - introducing factors (for age)
  - arrange
  - summarise
    - `n()`
  - count
  - group_by
- handling data with base-R
- tidy data
  - definition
  - table1 to table4
  - advantages
    - 1. uniformity, learning tools
    - vectorisation
  - reasons for untidy data
    - data entry
  - https://tidyr.tidyverse.org/
  - `billboard` data
- the pipe
  - motivate combination of steps
  - show pipe
  - shortcut

Day 3


  
Day 4

- Probability
  - Motivation: Racing game
  - Friend got first place 6 out of 8
    races against you.
    Is she the better player?
  - Introducing hypothesis tests 
  - We are never testing what we want to known, the inverse logic of p-values
  - pbinom, binomial probability
  




Day 5



Day 6

- Spotify data
- tidymodels
- PCA

Day 7


Day 8


Unassigned

- graphics devices, saving plots
-

# Exercises

# Solutions

# Resources
