---
title: "Lecture 8"
description: |
  Lecture 8: Into the Unknown — Exploring new Data
draft: false
author:
  - name: Jannik Buhr 
    url: https://jmbuhr.de
date: 12-20-2020
output:
  distill::distill_article:
    self_contained: false
    toc: true
categories:
  - lecture
bibliography: ["../../references.bib"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, warning = FALSE, message = FALSE
)
```


In today's lecture, the script is the result of experimenting
with unseen data as seen in the video. You can refer back
to it in order to get the code if the video goes to fast.
By itself, the script will be less useful than the script
for regular lectures. I hope you can learn something from
how I approach a new task and how I handle mistakes and
errors. The video is largely unedited; I only removed a
portion where all attempts at webscraping failed for a while
because my internet connection was gone...

# Video

Watch today's video here:

```{=html}
<iframe height="400" src="https://www.youtube.com/embed/E-pY4da9WCc" frameborder="0" allow="accelerometer; autoplay; clipboard-write;
encrypted-media; gyroscope; picture-in-picture" allowfullscreen>
</iframe>
```

# Slides

This special Christmas video has no slides.

# Script

```{r}
library(tidyverse)
library(rvest)
library(glue)
```

## Christmas Theme!

-   Christmas RStudio theme: <https://github.com/gadenbuie/rsthemes>

## Advanced Rmarkdown

-   Using the RStudio plugin to insert citations directly from Zotero
    -   preview of
        <https://blog.rstudio.com/2020/11/09/rstudio-1-4-preview-citations/>
-   Other output formats:
    -   <https://bookdown.org/yihui/rmarkdown/>
    -   <https://github.com/rstudio/rticles>
        -   note: tinytex package for pdf outputs

## Finding Help Easier

-   <https://reprex.tidyverse.org/>

## Into the Unknown: An Example Analysis of Unseen Data

![Into the Unknown, from the giphy
api](https://media0.giphy.com/media/m8WyNwXQUtzuijrtFq/giphy-downsized-medium.gif){.external}

### The Data: Fast Food Entrees

Find the data
[here](https://github.com/rfordatascience/tidytuesday/tree/master/data/2018/2018-09-04).

```{r}
fastfood <- read_csv("data/fastfood.csv") %>% 
  select(-salad, -X1) %>%  
  distinct(restaurant, item, .keep_all = TRUE)

fastfood
```

Watch the video to find out how these small data cleaning steps
came about.

```{r}
fastfood %>% 
  count(restaurant)
```

Counting things is always a good sanity check for your data.

```{r}
fastfood %>% 
  count(restaurant, item, sort = TRUE)
```

We will need the names of the restaurants.

```{r}
restaurants <- unique(fastfood$restaurant)
restaurants
```

This is not a very refined plot. I really wanted to get to
to the webscraping part to show you new exciting things.

```{r}
fastfood %>% 
  filter(restaurant == restaurants[2]) %>% 
  pivot_longer(-c(restaurant, item)) %>% 
  ggplot(aes(value, color = restaurant)) +
  geom_density() +
  facet_wrap(~ name, scales = "free")
```

### Webscraping with `rvest`

-   <https://rvest.tidyverse.org/>

The restaurant names on the links to
the complaints-website are slightly different to
the names we have in our fastfood dataset.
Some things are just faster to do manually
than figuring out an algorithmic way of achieving.
Although in this case, it could have been easier
because there is only one exception (sonic) to
"make everything lowercase and replace space with hyphens".

```{r}
restaurant_names <- c(
  "mcdonalds",
  "chick-fil-a",
  "sonic-drive-in",
  "arbys",
  "burger-king",
  "dairy-queen",
  "subway",
  "taco-bell"
)
```

Get some data! Try it on one example, then write a function
to do so. Then test the function on some examples.

```{r}
n <- 1
restaurant <- restaurant_names[8]
url <- glue("https://www.complaintsboard.com/{restaurant}/page/{1}")

get_complaints <- function(url) {
  read_html(url) %>% 
  html_nodes(css = ".list-text") %>% 
  html_text()
}
```

This chunk is set to `eval=FALSE`. Meaning we can still
run it manually but it will not run every time we knit the
document or press "run all chunks".
This enables us to keep the code for a long running computation
in the document while not having to run it every time.
A more elegant approach would have been to refactor this
into it's own R script after the initial exploration.

```{r, eval=FALSE}
complaints <- crossing(
  restaurant = restaurant_names,
  n = 1:30
) %>% 
  mutate(
    url = glue("https://www.complaintsboard.com/{restaurant}/page/{n}"),
    complaints = map(url, possibly(get_complaints, ""))
  )

write_rds(complaints, "data/complaints.rds")
```

Now we can load the scraped data.

```{r}
complaints <- read_rds("data/complaints.rds")
```

The is a lot of steps happening here.
First, we use the `tidytext` package to extract
words from the complaints, then we remove very
common words called stop words. We
then get the most commong words per restaurant
and finally match the original restaurant names
with the names we had to use for the webscarping links.

```{r}
names <- tibble(
  restaurants,
  restaurant_names
)

top_complaints <- complaints %>% 
  unnest_longer(complaints) %>% 
  tidytext::unnest_tokens(input = complaints, output = "word", token = "words") %>%
  anti_join(tidytext::stop_words, by = c("word")) %>% 
  count(restaurant, word, sort = TRUE) %>%
  group_by(restaurant) %>% 
  slice_max(order_by = n, n = 15) %>% 
  ungroup() %>% 
  left_join(names, by = c("restaurant" = "restaurant_names")) %>% 
  select(-restaurant) %>% 
  rename(restaurant = restaurants)

top_complaints %>% 
  rmarkdown::paged_table()
```

Next, we do a bunch of copy-pasting to quickly prototype an
app, where the user can select a restaurant, get
a quick overview of the feature distributions
and see the most common words in complaints.

## A shiny app

<https://shiny.rstudio.com/>

See `app.R` in the folder for lecture 8.
Note, that because shiny apps are usually in
their own project, not in some nested
file structure of a course website,
there are some shenanigans going on with
the file paths. When run locally,
R still has the working directory of the
project. But when deployed, only
the folder in which the app.R file lives
is deployed, with no notion of being
at the end of some project file path.
So the folder where the app.R file lives
becomes the new file path when the app is deployed
[online](https://buhr.shinyapps.io/lecture8/).

# Feedback

I will send round a link with a feedback form.
It is anonymous, so I have no way of tracking who 
submitted it. I will just assume you all did and
count it as a completed exercise.
