---
title: "Lecture 3"
description: |
  TODO
draft: true
author:
  - name: Jannik Buhr 
    url: https://jmbuhr.de
date: 11-02-2020
output:
  distill::distill_article:
    self_contained: false
    toc: true
categories:
  - lecture
bibliography: ["../../references.bib"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

# Video

Watch today's video here:

<iframe width="560" height="315" src="https://www.youtube.com/embed/AgpWX18dby4" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

# Slides

Hover over the slides and press **f** for full screen.
Press **?** for a list of keyboard shortcuts.

```{r}
knitr::include_url("slides2.html")
```

# Script

## Tidy Data

> "Happy families are all alike;
  every unhappy family is unhappy in its own way."
  $-$ Leo Tolstoy

> Tidy datasets are all alike,
  but every messy dataset is messy in its own way."
  $-$ Hadley Wickham

todo why tidy data?

### Making Data Tidy

with the `tidyr` package.

### Data that is better in another Format

e.g. matrices
todo explain matrix, show matrix maths,
briefly touch on arrays


# Exercises

# Solutions

# Resources
